Notes

### 2022-09-02 chapters/appendix-code.tex
Finish document

\begin{outcomment}
  Comentar cómo usar el código, las versiones de los paquetes, y los notebooks.

  Contar que son clases a la sklearn y compatibles con ska-fda. Enlazar al github.

  El proceso es reproducible (semilla 2022)

  Se usan 4 cores.

\end{outcomment}

>>>>

The Python source code developed for this work is available at the GitHub repository \url{https://github.com/antcc/rk-bfr}. The code is adequately documented and is structured in several directories as follows:

\begin{itemize}
  \item In the \texttt{rkbfr} folder we find the files responsible for the implementation of our models, separated according to the functionality they provide.
  \item The \texttt{reference\_methods} folder contains the implementation of the functional comparison algorithms that were not available through a standard Python library.
  \item At the root folder we have files for executing our experiments, which accept many user-specified parameters (e.g. number of iterations, type of data set, \ldots), along with some utility files. In particular, the script \texttt{results\_cv.py} contains the code for our comparison experiments, while script the \texttt{results\_all.py} allows the execution of our models without a cross-validation loop.
\end{itemize}

When possible, the code was implemented in a generic way that would allow for easy extensions or derivations, and also with efficiency in mind, so many functions and methods exploit the vectorization capabilities of the \textit{numpy} and \textit{scipy} libraries. Moreover, we followed closely the style of the \textit{scikit-learn} and \textit{scikit-fda} libraries, so our methods are fully compatible and could be integrated (after some minor tweaking) with both of them. The code for the experiments was executed in a machine with 4 cores, and a random seed with value 2022 was set for reproducibility. We provide a script file \texttt{launch.sh} that illustrates a typical execution. Lastly, there are also Jupyter notebooks that demonstrate the use of our methods in a more visual way. Inside these notebooks there is a step-by-step guide on how one might execute our algorithms, accompanied by many graphical representations, and offering the possibility of changing multiple parameters to experiment with the code. In addition, there is also a notebook that can be used to generate all the tables and figures of this document pertaining to the experimental results.

### 2022-09-02 chapters/background-related-work.tex
Finish document

 The most common scalar-on-function linear regression model is the classical \(L^2\)-model, widely popularized since the first edition (1997) of the monograph by~\citet{ramsay2005functional}. It can be seen as a generalization of the usual finite-dimensional model, replacing the scalar product in \(\R^p\) for that of the functional space \(L^2[0,1]\):
\begin{equation}\label{eq:l2-linear-model}
Y = \alpha_0 + \dotprod{X}{\beta} + \epsilon = \alpha_0 + \int_0^1 X(t)\beta(t)\, dt + \epsilon,
\end{equation}
where \(\alpha_0\in \R\), \(\epsilon\) is a random error term independent from \(X\) with \(\E [\epsilon]=0\) and \(\E[X(t)\epsilon]=0\) almost everywhere; and the functional slope parameter \(\beta=\beta(\cdot)\) is assumed to be a member of the infinite-dimensional space \(L^2[0, 1]\). A careful rearrangement of~\eqref{eq:l2-linear-model} shows that the model is equivalently expressed as \(\Delta=\K\beta\), where \(\Delta(\cdot)\) is the cross-covariance function of \(X\) and \(Y\), and \(\K\) is the covariance operator of the process \(X\) (to be defined in Section~\ref{sec:rkhs}). This expression is a continuous analog of the normal equations that arise in finite-dimensional regression settings, and since the operator \(\K\) is non-invertible in general, questions of existence and unicity of \(\beta\) need further study (see \citet{cardot2011functional} for more insights).

@@

In spite of the apparent generality of model~\eqref{eq:l2-linear-model}, it can be shown that it is not flexible enough to include ``simple'' finite-dimensional models based on linear combinations of the marginals, such as \(Y=\alpha_0 + \beta_1 X(t_1)+ \cdots + \beta_p X(t_p) + \epsilon\) for some constants \(\beta_j\in\R\) and instants \(t_j\in[0,1]\); see \citet{berrendero2020general} for additional details on this. It turns out that in both linear and logistic scenarios a natural alternative to the \(L^2\)-model is the so-called Reproducing Kernel Hilbert Space (RKHS) model, which instead assumes the unknown functional parameter to be a member of the RKHS associated with the covariance function of the process \(X\), making use of the scalar product of that functional space. As we will show later on, not only is this model simpler and arguably easier to interpret, but it also constrains the parameter space to smoother and more manageable functions. In fact, it does include a model based on finite linear combinations of the marginals of \(X\) as a particular case, which is especially appealing to practitioners confronted with functional data problems due to its simplicity. These RKHS-based models and their idiosyncrasies have been explored in \citet{berrendero2019rkhs, berrendero2020general} in the functional linear regression setting, and in \citet{berrendero2018functional, berrendero2018use} for the case of functional logistic regression.

There are other proposals for models that change the habitat of the functional parameter, and there are even some in which \(\beta\) is supposed to live in a RKHS, most notably \citet{yuan2010reproducing}. However, these are arbitrary RKHS's that do not directly exploit the relation with the process \(X\) (though the authors use said connection to derive some asymptotic properties of their estimators), and therefore the approach is somewhat different. Incidentally, the RKHS associated with \(X\) also has some interesting properties that contribute to shed light on the near-perfect classification phenomenon for functional data, described by \citet{delaigle2012achieving} and further examined for example in the works of \citet{berrendero2018use} or \citet{torrecilla2020optimal}.

@@

\begin{proposition} The following properties hold for the space \(\Hcal(K)\):
  \begin{enumerate}
    \item The evaluation operator \(\delta_t\) is bounded for all \(t\in[0,1]\).
    \item Norm convergence implies pointwise convergence, and if \(K\) is continuous, it also implies uniform convergence.
    \item If \(K\) is m-times continuously differentiable, then so is every function \(f\in\Hcal(K)\). In particular, if \(K\) is continuous, every function \(f\in\Hcal(K)\) is continuous.
  \end{enumerate}
\end{proposition}
\begin{proof}
  To prove \((i)\), simply note that for any \(t\in[0,1]\) and \(f\in\Hcal(K)\) the Cauchy-Schwarz inequality and the reproducing property tell us that
  \[
  |\delta_t(f)| = |f(t)| = |\dotprod{f}{K(t, \cdot)}_K| \leq \|K(t, \cdot)\|_K \|f\|_K = \sqrt{K(t, t)}\|f\|_K,
  \]
  and consequently \(\|\delta_t\|\leq \sqrt{K(t, t)}\). To see \((ii)\), observe that for any \(t\in[0,1]\) we have
  \[
    |f_n(t) - f(t)| = |\delta_t(f_n - f)| \leq \|\delta_t\|\|f_n - f\|_K,
  \]
  and we have just shown above that the evaluation operator is bounded. Moreover, if \(K\) is continuous on its (compact) domain, \(\|\delta_t\| \leq \sup_s \sqrt{K(s, s)}=M<\infty\) for all \(t\), so the convergence is indeed uniform. Finally, we prove the second statement in \((iii)\), and refer the reader to \citet[][Th 2.6]{saitoh2016theory} for a complete proof. If \(K\) is continuous, then it is clear that every function in \(\Hcal_0(K)\) is continuous. Let us now fix an arbitrary \(f\in \Hcal(K)\). By definition, there exists a sequence \(\{f_n\}\subset \Hcal_0(K)\) of continuous functions such that \(\|f_n - f\|_K \to 0\). But then \(f_n \to f\) uniformly by \((ii)\), and therefore \(f\) is continuous as the uniform limit of continuous functions. Note that this statement remains valid, with a slightly different proof, even when the underlying domain is not compact.
\end{proof}

@@

\begin{theorem}[Mercer's theorem]\label{th:mercer}
    Let \(K\) be a continuous kernel on \([0,1]^2\). Then, there exists an orthonormal basis \(\{\phi_j\}\) of \(L^2[0,1]\) consisting of eigenfunctions of \(\K\), whose corresponding eigenvalues are all positive and form a decreasing sequence \(\{\lambda_j\}\to 0\), and such that
    \[
      K(t, s) = \sum_{j=1}^\infty \lambda_j \phi_j(t)\phi_j(s)
    \]
    for all \(t\) and \(s\), where the convergence is absolute and uniform.
\end{theorem}

Furthermore, a similar orthogonal decomposition holds for our stochastic process \(X\), an expansion which lies at the base of many \(L^2\) functional regression models. The proof of the following result, along with the proof of Mercer's theorem, can be consulted in \citet[][Ch.~3.2]{berlinet2004reproducing}.

\begin{theorem}[Karhunen-Loève expansion]
  With the assumptions and notations of Theorem~\eqref{th:mercer}, the centered process \(X=X(t)\) with covariance function \(K\) admits the quadratic-mean representation
  \[
  X(t) = \sum_{j=1}^\infty \zeta_j \phi_j(t),
  \]
  where the convergence is in \(L^2(\Omega)\) and uniform in \(t\). The \(\zeta_j\) are independent zero-mean random variables with \(\E[\zeta_i\zeta_j]=\delta_{ij}\lambda_j\), given explicitly by the formula
  \[
    \zeta_j = \int_0^1 X(t)\phi_j(t)\, dt, \quad j\geq 1.
  \]
\end{theorem}

In addition to these results, it is worth mentioning that the covariance operator \(\K\) provides several alternative definitions of \(\Hcal(K)\). For example, the RKHS can be identified with the image of the operator's square root, i.e., \(\Hcal(K) = \K^{1/2}(L^2[0, 1])\), with inner product \(\dotprod{f}{g}_K = \dotprod{\K^{-1/2}(f)}{\K^{-1/2}(g)}\). Furthermore, we can think of the norm in \(\Hcal(K)\) as an \(L^2\)-like regularized norm, since this space can be seen as \(\Hcal(K) = \{f \in L^2[0, 1]: \ \sum_j \lambda_j^{-1}\dotprod{f}{\phi_j}^2 < \infty \}\), with corresponding inner product \(\dotprod{f}{g}_K = \sum_j \lambda_j^{-1}\dotprod{f}{\phi_j}\dotprod{g}{\phi_j}\). Note that, since \(\{\lambda_j\}\) tends to zero, this definition highlights the fact that functions in \(\Hcal(K)\) are smooth, not only in that they inherit the regularity of the kernel \(K\), but also in the sense that their components in an orthonormal basis, namely \(\dotprod{f}{\phi_j}\), need to vanish quickly.

@@

However, despite the close connection between the process \(X\) and the space \(\Hcal(K)\), special care must be taken when dealing with concrete realizations of the process. It can be shown that under rather general conditions the trajectories of \(X\) \textit{do not belong} to the corresponding RKHS with probability one \citep[see for example][Cor.~7.1]{lukic2001stochastic}. For instance, the trajectories of a Brownian motion are known to be nowhere differentiable, but as shown in Example~\ref{ex:bm}, the elements of the corresponding RKHS are differentiable almost everywhere. An heuristic argument to justify this fact, given in \citet{wahba1990spline}, is as follows: consider the Karhunen-Loève expansion of \(X\), i.e.,
\[
  X(t) = \sum_{j=1}^\infty \zeta_j \phi_j(t),
\]
and the truncated version \(X_N(t)\) up to the \(N\)-th term. On the one hand, for each fixed \(t\) we have \(X_N(t)\to X(t)\) in the quadratic mean sense (by the Karhunen-Loève theorem), but on the other hand observe that
\[
  \E\|X_N(\cdot)\|^2_K = \E \sum_{j=1}^N \frac{\zeta_j^2}{\lambda_j} = N \to \infty \quad (N\to\infty).
\]

@@

Generally speaking, a MCMC algorithm can be described at time \(t+1\) as \(\Lambda(t+1)=R(\Lambda(t), \xi(t), \pi)\), where \(\Lambda(t)\) is the state of the chain at instant \(t\), \(\pi(x)\) is the objective distribution, and \(\xi(t)\) is a sequence of i.i.d. random variables that represent the random behavior of the chain. With this notation, the property of affine-invariant can be characterized as \citep{goodman2010ensemble}:
 \[
 R(A\lambda+b, \xi(t), \pi_{A,b}) = AR(\lambda, \xi(t), \pi) + b,
 \]
 for all \(A,b\) and \(\lambda\), and for almost all \(\xi(t)\). This means that if we fix a random generator and run the algorithm twice, one time using \(\pi\) and starting in \(\Lambda(0)\) and a second time using \(\pi_{A,b}\) with initial point \(\Gamma(0)=A\Lambda(0)+b\), then \(\Gamma(t)=A\Gamma(t)+b\) for all \(t\). In \citep{goodman2010ensemble} the authors consider an ensemble of samplers with the affine invariance property. Specifically, they work with a set \(\Lambda=(\Lambda_1, \dots, \Lambda_L)\) of \textit{walkers}, where \(\Lambda_l(t)\) represents an individual chain at time \(t\). At each iteration, an affine-invariant transformation is used to find the next point, which is constructed using the current values of the rest of the walkers (similar to Gibb's algorithm), namely the \textit{complementary ensemble}
 \[
  \Lambda_{-l}(t) = \{\Lambda_1(t+1), \dots, \Lambda_{l-1}(t+1), \Lambda_{l+1}(t), \dots, \Lambda_L(t)\}, \quad l=1,\dots, L.
  \]

>>>>

 The most common scalar-on-function linear regression model is the classical \(L^2\)-model, widely popularized since the first edition (1997) of the monograph by~\citet{ramsay2005functional}. It can be seen as a generalization of the usual finite-dimensional model, replacing the scalar product in \(\R^p\) for that of the functional space \(L^2[0,1]\):
\begin{equation}\label{eq:l2-linear-model}
Y = \alpha_0 + \dotprod{X}{\beta} + \epsilon = \alpha_0 + \int_0^1 X(t)\beta(t)\, dt + \epsilon,
\end{equation}
where \(\alpha_0\in \R\), \(\epsilon\) is a random error term independent from \(X\) with \(\E [\epsilon]=0\), and the functional slope parameter \(\beta=\beta(\cdot)\) is assumed to be a member of the infinite-dimensional space \(L^2[0, 1]\). A careful rearrangement of~\eqref{eq:l2-linear-model} shows that the model is equivalently expressed as \(\Delta=\K\beta\), where \(\Delta(\cdot)\) is the cross-covariance function of \(X\) and \(Y\), and \(\K\) is the covariance operator of the process \(X\) (to be defined in Section~\ref{sec:rkhs}). This expression is a continuous analog of the normal equations that arise in finite-dimensional regression settings, and since the operator \(\K\) is non-invertible in general, questions of existence and unicity of \(\beta\) need further study (see \citet{cardot2011functional} for more insights).

@@

In spite of the apparent generality of model~\eqref{eq:l2-linear-model}, it can be shown that it is not flexible enough to include ``simple'' finite-dimensional models based on linear combinations of the marginals, such as \(Y=\alpha_0 + \beta_1 X(t_1)+ \cdots + \beta_p X(t_p) + \epsilon\) for some constants \(\beta_j\in\R\) and instants \(t_j\in[0,1]\); see \citet{berrendero2020general} for additional details on this. It turns out that in both linear and logistic scenarios a natural alternative to the \(L^2\)-model is the so-called Reproducing Kernel Hilbert Space (RKHS) model, which instead assumes the unknown functional parameter to be a member of the RKHS associated with the covariance function of the process \(X\), making use of the scalar product of that functional space. As we will show later on, not only is this model simpler and arguably easier to interpret, but it also constrains the parameter space to smoother and more manageable functions. In fact, it does include a model based on finite linear combinations of the marginals of \(X\) as a particular case, which is especially appealing to practitioners confronted with functional data problems due to its simplicity. These RKHS-based models and their idiosyncrasies have been explored in \citet{berrendero2019rkhs, berrendero2020general} in the functional linear regression setting, and in \citet{berrendero2018functional} for the case of functional logistic regression.

There are other proposals for models that change the habitat of the functional parameter, and there are even some in which \(\beta\) is supposed to live in a RKHS, most notably \citet{yuan2010reproducing}. However, theirs are arbitrary RKHS's that do not directly exploit the relation with the process \(X\) (though the authors do use this connection to derive some asymptotic properties of their estimators), and therefore the approach is somewhat different. Incidentally, the RKHS associated with \(X\) also has some interesting properties that contribute to shed light on the near-perfect classification phenomenon for functional data, described by \citet{delaigle2012achieving} and further examined for example in the works of \citet{berrendero2018use} or \citet{torrecilla2020optimal}.

@@

\begin{proposition} The following properties hold for the space \(\Hcal(K)\):
  \begin{enumerate}
    \item The evaluation operator \(\delta_t\) is bounded for all \(t\in[0,1]\).
    \item Norm convergence implies pointwise convergence, and if \(K\) is continuous, it also implies uniform convergence.
    \item If \(K\) is m-times continuously differentiable, then so is every function \(f\in\Hcal(K)\). In particular, if \(K\) is continuous, every function \(f\in\Hcal(K)\) is continuous.
  \end{enumerate}
\end{proposition}
\begin{proof}
  To prove \((i)\), simply note that for any \(t\in[0,1]\) and \(f\in\Hcal(K)\) the Cauchy-Schwarz inequality and the reproducing property tell us that
  \[
  |\delta_t(f)| = |f(t)| = |\dotprod{f}{K(t, \cdot)}_K| \leq \|K(t, \cdot)\|_K \|f\|_K = \sqrt{K(t, t)}\|f\|_K,
  \]
  and consequently \(\|\delta_t\|\leq \sqrt{K(t, t)}\). To see \((ii)\), observe that for any \(t\in[0,1]\) we have
  \[
    |f_n(t) - f(t)| = |\delta_t(f_n - f)| \leq \|\delta_t\|\|f_n - f\|_K,
  \]
  as we have just shown in \((i)\) that the evaluation operator is bounded. Moreover, if \(K\) is continuous on its (compact) domain, \(\|\delta_t\| \leq \sup_s \sqrt{K(s, s)}=M<\infty\) for all \(t\), so the convergence is indeed uniform. Finally, we prove the second statement in \((iii)\), and refer the reader to \citet[][Th 2.6]{saitoh2016theory} for a complete proof. If \(K\) is continuous, then it is clear that every function in \(\Hcal_0(K)\) is continuous. Let us now fix an arbitrary \(f\in \Hcal(K)\). By definition, there exists a sequence \(\{f_n\}\subset \Hcal_0(K)\) of continuous functions such that \(\|f_n - f\|_K \to 0\). But then \(f_n \to f\) uniformly by \((ii)\), and therefore \(f\) is continuous as the uniform limit of continuous functions. Note that this statement remains valid, with a slightly different proof, even when the underlying domain is not compact.
\end{proof}

@@

\begin{theorem}[Mercer's theorem]\label{th:mercer}
    Let \(K\) be a continuous kernel on \([0,1]^2\). Then, there exists an orthonormal system \(\{\phi_j\}\) in \(L^2[0,1]\) consisting of eigenfunctions of \(\K\), whose corresponding eigenvalues are all positive and form a non-increasing sequence \(\{\lambda_j\}\to 0\), and such that
    \[
      K(t, s) = \sum_{j=1}^\infty \lambda_j \phi_j(t)\phi_j(s)
    \]
    for all \(t\) and \(s\), where the convergence is absolute and uniform.
\end{theorem}

In connection with our RKHS theory, it can be shown that the set \(\{\sqrt{\lambda_j}\phi_j\}\) constitutes an orthonormal basis of \(\Hcal(K)\) \citep[see e.g.][Sec.~4.4]{cucker2007learning}. Furthermore, a similar orthogonal decomposition holds for our stochastic process \(X\), an expansion which lies at the base of many \(L^2\) functional regression models. The proof of the following result, along with the proof of Mercer's theorem, can be consulted in \citet[][Sec.~3.2]{berlinet2004reproducing}.

\begin{theorem}[Karhunen-Loève expansion]
  With the assumptions and notations of Theorem~\ref{th:mercer}, the centered process \(X=X(t)\) with covariance function \(K\) admits the quadratic-mean representation
  \[
  X(t) = \sum_{j=1}^\infty \zeta_j \phi_j(t),
  \]
  where the convergence is in \(L^2(\Omega)\) and uniform in \(t\). The \(\zeta_j\) are independent zero-mean random variables with \(\E[\zeta_i\zeta_j]=\delta_{ij}\lambda_j\), given explicitly by the formula
  \[
    \zeta_j = \int_0^1 X(t)\phi_j(t)\, dt, \quad j\geq 1.
  \]
\end{theorem}

In addition to these results, it is worth mentioning that the covariance operator \(\K\) provides several alternative definitions of \(\Hcal(K)\). For example, the RKHS can be identified with the image of the operator's square root, i.e., \(\Hcal(K) = \K^{1/2}(L^2[0, 1])\), with inner product \(\dotprod{f}{g}_K = \dotprod{\K^{-1/2}(f)}{\K^{-1/2}(g)}\). Furthermore, we can think of the norm in \(\Hcal(K)\) as an \(L^2\)-like regularized norm, since this space can also be seen as the set \(\Hcal(K) = \{f \in L^2[0, 1]: \ \sum_j \lambda_j^{-1}\dotprod{f}{\phi_j}^2 < \infty \}\), with corresponding inner product \(\dotprod{f}{g}_K = \sum_j \lambda_j^{-1}\dotprod{f}{\phi_j}\dotprod{g}{\phi_j}\). Note that, since \(\{\lambda_j\}\) tends to zero, this definition highlights the fact that functions in \(\Hcal(K)\) are smooth, not only in that they inherit the regularity of the kernel \(K\), but also in the sense that their components in an orthonormal basis, namely \(\dotprod{f}{\phi_j}\), need to vanish quickly.

@@

However, despite the close connection between the process \(X\) and the space \(\Hcal(K)\), special care must be taken when dealing with concrete realizations of the process. It can be shown that under rather general conditions the trajectories of \(X\) \textit{do not belong} to the corresponding RKHS with probability one \citep[see for example][Cor.~7.1]{lukic2001stochastic}. For instance, the trajectories of a Brownian motion are known to be nowhere differentiable, but as shown in Example~\ref{ex:bm}, the elements of the corresponding RKHS are differentiable almost everywhere. An heuristic argument to justify this fact, given in \citet{wahba1990spline}, is as follows: consider the Karhunen-Loève expansion of \(X\), i.e.,
\[
  X(t) = \sum_{j=1}^\infty \zeta_j \phi_j(t),
\]
and the truncated version \(X_N(t)\) up to the \(N\)-th term. On the one hand, for each fixed \(t\) we have \(X_N(t)\to X(t)\) in the quadratic mean sense (by the Karhunen-Loève theorem), but on the other hand observe that
\[
  \E\left[\|X_N(\cdot)\|^2_K\right] = \E \left[\sum_{j=1}^N \frac{\zeta_j^2}{\lambda_j}\right] = N \to \infty \quad (N\to\infty).
\]

@@

Generally speaking, a MCMC algorithm can be described at time \(t+1\) as \(\Lambda(t+1)=R(\Lambda(t), \xi(t), \pi)\), where \(\Lambda(t)\) is the state of the chain at instant \(t\), \(\pi(x)\) is the objective distribution, and \(\xi(t)\) is a sequence of i.i.d. random variables that represent the random behavior of the chain. With this notation, the property of affine-invariant can be characterized as \citep{goodman2010ensemble}:
 \[
 R(A\lambda+b, \xi(t), \pi_{A,b}) = AR(\lambda, \xi(t), \pi) + b,
 \]
 for all \(A,b\) and \(\lambda\), and for almost all \(\xi(t)\). This means that if we fix a random generator and run the algorithm twice, one time using \(\pi\) and starting in \(\Lambda(0)\) and a second time using \(\pi_{A,b}\) with initial point \(\Gamma(0)=A\Lambda(0)+b\), then \(\Gamma(t)=A\Lambda(t)+b\) for all \(t\). In \citet{goodman2010ensemble} the authors consider an ensemble of samplers with the affine invariance property. Specifically, they work with a set \(\Lambda=(\Lambda_1, \dots, \Lambda_L)\) of \textit{walkers}, where \(\Lambda_l(t)\) represents an individual chain at time \(t\). At each iteration, an affine-invariant transformation is used to find the next point, which is constructed using the current values of the rest of the walkers (similar to Gibb's algorithm), namely the \textit{complementary ensemble}
 \[
  \Lambda_{-l}(t) = \{\Lambda_1(t+1), \dots, \Lambda_{l-1}(t+1), \Lambda_{l+1}(t), \dots, \Lambda_L(t)\}, \quad l=1,\dots, L.
  \]

### 2022-09-02 chapters/bayesian-functional-regression.tex
Finish document

As far as prior distributions go, we propose to use the same ones as we did in~\eqref{eq:prior-linear} for the linear regression model. However, in this case the parameter \(\sigma^2\) is a nuisance parameter and can be safely ignored at the end, since it does not appear in the final model. The posterior distribution is then derived after a routine calculation.

>>>>

As far as prior distributions go, we propose to use the same ones as we did in~\eqref{eq:prior-linear} for the linear regression model. However, in this case the nuisance parameter \(\sigma^2\) only appears as part of the hierarchical prior distribution, and not in the final model. The posterior distribution is then derived after a routine calculation.

### 2022-09-02 chapters/conclusions.tex
Finish document

\begin{outcomment}
  - Destacar también las ventajas de lo bayesiano: flexibilidad, incorporación de información con distintas priors, etc. En la introducción, conclusiones, experimentos, ..
- Destacar interpretabilidad en los experimentos: e.g. un valor de beta=3.4 se puede interpretar fácilmente en un caso concreto. En general, nuestros modelos tienen menor dimensión.
- Logistic un poco mejor que linear?
- Tiempo de ejecución alto (característico de los algoritmos MCMC), a cambio mucha sencillez.
- CV ha sido pequeño. Con más parámetros, probablemente mejores resultados. También flexibilidad para cambiar las prioris.
- We do not advise to use the mean as a summary statistic.
\end{outcomment}

@@

\begin{outcomment}
  - Mejorar pymc.
  - consistencia de la posteriori, robustez de la priori
  - Otras prioris

\end{outcomment}

>>>>

In this work we have addressed some inference and prediction issues involved in the task of learning from functional data. In particular, we have tackled functional linear and logistic regression problems, and we have presented an alternative to the \(L^2\)-models predominant in these scenarios. Our proposed models rely heavily on the theory of RKHS's, offering finite-dimensional approximations from a functional perspective that are conceptually simpler and facilitate the interpretation and subsequent analysis, an attribute that practitioners surely find especially compelling. The underlying theoretical aspects are nevertheless far from trivial, so we have also explored some of the related mathematical background in this text.

With the aid of Bayesian methods, we have introduced a computationally feasible approach to parameter estimation within the proposed RKHS models. The flexibility is improved through the use of prior distributions that can introduce pre-existing information into the model. In addition, the fact that we have a complete posterior distribution at our disposal enables us to derive different prediction and variable selection methods, or even perform other types of inference such as the construction of credible intervals. Moreover, the use of MCMC methods to approximate the posterior distribution results in an implementation that is simpler than most existing alternatives. Although this simplicity comes at the expense of increased computational requirements, this computational cost is still reasonable and the model can be used in practice.

We implemented the whole inference and prediction pipeline in the Python programming language, and then used it to carry out a large empirical study to assess the performance of our models against other usual alternatives, functional and otherwise. The results were generally satisfactory, proving that we have competitive methods that achieve good results in different scenarios, often with low dimensionality and thus increased tractability and interpretability. Lastly, we believe that with more computational resources we could consider a larger pool of hyperparameters in the model selection phase, which would most likely contribute to better results.

@@

Some lines of further research suggested by our work are the following.

\begin{itemize}
  \item To delve further into the relationship between RKHS's and functional data problems, a connection that has proven to be fruitful in many scenarios. A first idea would be to extend the RKHS-based logistic regression model to a generalized functional linear model with an arbitrary link function.
  \item To try to derive some theoretical properties of our Bayesian predictors. For example, consistency and/or robustness results regarding the posterior distribution would be an excellent complement to the practical side of this work.
  \item To find other prior distributions for our parameters that perform better in general, or to eliminate the need of the tuning hyperparameter \(\eta\).
  \item To experiment with other MCMC algorithms for posterior approximation. Specifically, it would be very interesting to implement an efficient and reliable reversible-jump MCMC method in Python, which as we have already mentioned is a better fit for our particular Bayesian model.
\end{itemize}

### 2022-09-02 chapters/experiments.tex
Finish document

\paragraph{Logistic regression data sets.} Again we consider three different types of data sets,  with a common value of \(\alpha_0=-0.5\). In this case we randomly permute 10\% of the labels to introduce some noise in the simulations.
\begin{itemize}
  \item Four logistic RKHS models with the same functional parameter as in the linear regression case (one for each GP). Specifically,
  \[
  \P(Y=1\mid X) = \frac{1}{e^{0.5 +5X(0.1) - X(0.4) - 10X(0.8)}}.
  \]
  \item Four logistic \(L^2\)-models with the same coefficient function as in the linear regression case.
  \item A ``mixture'' model based on non-GP regressors, in which we mix regressors from two different GP's and label them according to their origin. Firstly, we consider a homoscedastic case to distinguish between a standard Brownian motion and a Brownian motion with a mean function that is zero until \(t=0.5\), and then becomes \(m(t)=0.75t\). Secondly, we consider a heteroscedastic case to distinguish between a standard Brownian motion and a Brownian motion with variance \(2\), that is, with kernel \(K(t,s)=2\min\{t,s\}\).
\end{itemize}
In this case there are again three real data sets. The first one is a subset of the Medflies data set \citep{carey1998relationship}, consisting on samples of the number of eggs laid daily by 534 flies over 30 days, trying to predict if their longevity is high or low. The second one is the Berkeley Growth Study data set \citep{tuddenham1954physical}, which records the height of 54 girls and 39 boys over 31 different points in their lives. Finally, we have selected a subset of the Phoneme data set \citep{hastie1995penalized}, based on 200 digitized speech frames over 128 equispaced points to predict the phonemes ``aa'' and ``ao''.

@@

The main parameters of all these algorithms are selected by cross-validation, using the same 5 folds as our proposed models so that the comparisons are fair. In particular, the regularization parameter is searched among 20 values in the logarithmic space \([10^{-4}, 10^4]\), the number of manually selected variables is one of \(\{5, 10, 15, 20, 25, 50\}\), the number of components for dimension reduction and variable selection techniques is in the set \(\{2, 3, 4, 5, 7, 10, 15, 20\}\), the number of basis elements for quadratic spline bases is in \(\{8,10,12,14,16\}\), the number of basis elements for Fourier bases is in \(\{3,5,7,9,11\}\), and the number of neighbors in the KNN classifier is in \(\{3,5,7,11\}\).

@@

Finally, Figure~\ref{fig:clf_emcee_nongp} shows that our classifiers perform better than most comparison algorithms when separating two homoscedastic processes, but they struggle in the heteroscedastic case. Our variable selection methods are close to the global mean accuracy in both cases.

@@

One requirement that our model should satisfy is that it be able to recover the true parameter function when the underlying model is a RKHS model. This is generally the case when the value of \(p\) in our model and the true value of \(p\) coincide, but what happens when we change the value of \(p\) in the model?

>>>>

\paragraph{Logistic regression data sets.} Again we consider three different types of data sets,  with a common value of \(\alpha_0=-0.5\). In this case we randomly permute 10\% of the labels to introduce some noise in the simulations.
\begin{itemize}
  \item Four logistic RKHS models with the same functional parameter as in the linear regression case (one for each GP). Specifically,
  \[
  \P(Y=1\mid X) = \frac{1}{1 + e^{0.5 +5X(0.1) - X(0.4) - 10X(0.8)}}.
  \]
  \item Four logistic \(L^2\)-models with the same coefficient function as in the linear regression case.
  \item A ``mixture'' model based on non-GP regressors, in which we mix regressors from two different GP's (with equal probability \(p=1/2\)) and label them according to their origin. Firstly, we consider a homoscedastic case to distinguish between a standard Brownian motion and a Brownian motion with a mean function that is zero until \(t=0.5\), and then becomes \(m(t)=0.75t\). Secondly, we consider a heteroscedastic case to distinguish between a standard Brownian motion and a Brownian motion with variance \(2\), that is, with kernel \(K(t,s)=2\min\{t,s\}\).
\end{itemize}
In this case there are again three real data sets. The first one is a subset of the Medflies data set \citep{carey1998relationship}, consisting on samples of the number of eggs laid daily by 534 flies over 30 days, trying to predict if their longevity is high or low. The second one is the Berkeley Growth Study data set \citep{tuddenham1954physical}, which records the height of 54 girls and 39 boys over 31 different points in their lives. Finally, we have selected a subset of the Phoneme data set \citep{hastie1995penalized}, based on 200 digitized speech frames over 128 equispaced points to predict the phonemes ``aa'' and ``ao''.

@@

The main parameters of all these algorithms are selected by cross-validation, using the same 5 folds as our proposed models so that the comparisons are fair. In particular, the regularization parameter is searched among 20 values in the logarithmic space \([10^{-4}, 10^4]\), the number of manually selected variables is one of \(\{5, 10, 15, 20, 25, 50\}\), the number of components for dimension reduction and variable selection techniques is in the set \(\{2, 3, 4, 5, 7, 10, 15, 20\}\), the number of basis elements for quadratic spline bases is in \(\{8,10,12,14,16\}\), the number of basis elements for Fourier bases is one of \(\{3,5,7,9,11\}\), and the number of neighbors in the KNN classifier is in \(\{3,5,7,9,11\}\).

@@

Finally, Figure~\ref{fig:clf_emcee_nongp} shows that our classifiers perform better than most comparison algorithms when separating two homoscedastic processes, but they struggle in the heteroscedastic case. Incidentally, this heteroscedastic case of two zero-mean Brownian motions has a special interest, since it can be shown that the Bayes error is zero in the limit of dense monitoring (i.e. with an arbitrarily fine measurement grid), a manifestation of the ``near-perfect'' classification phenomenon analyzed for example in \citet{torrecilla2020optimal}. Our results are in line with the empirical studies of said article, where the authors conclude that even though the asymptotic theoretical error is zero, most methods are suboptimal in practice (possibly due to the high collinearity of the data), with the notable exception of PCA+QDA.

@@

One requirement that our model should satisfy is that it should be able to recover the true parameter function when the underlying model is a RKHS model. This is generally the case when the value of \(p\) in our model and the true value of \(p\) coincide, but what happens when we change the value of \(p\) in the model?
